Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-17 manuscript/bef_stability/manuscript.Rmd
Decrease number of words lower than 4980

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in foodwebs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can in some cases reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, recent models revealed new mechanisms by
which foodweb structure affects total ecosystem biomass and its temporal
variability. In particular, the number of trophic levels in foodwebs, which is
associated with the presence of species at higher trophic levels, increases
total primary production [@wang_biodiversity_2018]. The addition of predators has
also cascading effects on both species variability and synchrony at lower
trophic levels [@teng_dynamics_2004; @shanafelt_stability_2018].
Overall, these theoretical studies show that the effects of species richness and
network structure, such as connectance and average trophic level, are
potentially conflicting or synergistic in such a way that they need to be
addressed jointly.

@@

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing
data from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allows for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). For streams
shallower than 0.7m across the entire width, the sampling took place over the
whole width of the streams and was done by foot whereas in streams deeper than
0.7m, the sampling was done over the bank by foot or by boat. The sampling
season spanned from late spring to autumn. One sampling event corresponds to a
sampling a given year in a given sampling station. During sampling, all the
fishes caught were identified. The fishes were released at the end of
the sampling. Additionally, the width and the depth of the stream were 
recorded during sampling event as well as the sampling effort, quantified as the
surface sampled (range `r range_surface[1]`-`r range_surface[2]` square metres).

The stations where the stream was of intermediate size could be sampled
alternatively with one of the two sampling protocols. In this case, we
kept only the most frequently used sampling protocol. Then, we selected the
stations that had been sampled ten years or more. There was low heterogeneity of
the sampling month within stations, with the median sampled month being
mid-August while the median of the standard deviation of the sampled month
within station was $0.7$. As the concept of community variability referred to
variability around an equilibrium, which is rarely observed or difficult to
characterize in natural communities, only stations with no temporal trends
in total biomass were selected (detailed in appendix B, part 2). However, the
results were robust to the relaxation of this hypothesis (Figure S7). These
selection steps resulted in 99 stations distributed over seven hydrographic
basins (Figure \@ref(fig:matmet) A).

@@

The body size of all fish individuals was not recorded when they were too
numerous (Figure S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution whose mean and standard deviation were quantified from the
recorded individuals of the sampling (Figure S1, appendix B, part 1.2). The body
mass of each individual was then estimated following the results of two
meta-analyses on fishes [@lleonart_removing_2000; @froese_cube_2006], using an
allometric law $B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams,
$l$ the body size in millimeters, $i$ the individual and where the coefficients
$0.01$ and $3.03$.  Species biomass and total biomass at each sampling event
were then calculated by summing individual body masses. To account for the
variation in sampling effort among sites, species biomass and total biomass were
corrected by the surface sampled, and expressed in grams by square meter ($gm^{-2}$). In the analysis,
the temporal average of the annual total biomass of a station has been defined
as the median of the total biomass across sampling years.

@@

Species richness was computed as the total number of species present across
all the sampling events in a given station. As for biomass, we controlled
species richness for sampling effort by dividing the total number of species by
the total surface sampled in the given station. Species richness was then
expressed in number of species per square meter sampled.

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Figure
\@ref(fig:matmet) B), then (2) determine network of a given community by
extracting the metaweb according to the trophic species present at a given
site and sampling event (Figure \@ref(fig:matmet) C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Figure S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Figure \@ref(fig:matmet) B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. Species identity allowed to
differentiate between piscivorous and non piscivorous species, to know which
resources the trophic species are feeding on and to know the minimum size at
which a piscivorous species begin to feed on fishes. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by a body size
range (Figure S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Figure S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic specie was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Figure S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class, for both the piscivorous and non piscivorous trophic species.
Lastly, the trophic links among resource nodes were set according to the
literature [@allan_stream_2007; @hart_handbook_2008].

@@

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Figure S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Figure S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, DBO and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure as measured by connectance and
average trophic level [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Figure
\@ref(fig:sem-plot) A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Figure
\@ref(fig:sem-plot) C). Instead of the annual total biomass, we modelled the
synchrony and population variability which in turn determined the temporal
stability of biomass. The structural equation models used for annual total
biomass and its temporal stability contained respectively 4 and 6 response
variables, 35 and 26 direct paths.

The structural equation models were based on Gaussian linear models. All variables
except the connectance and average trophic level
were logged to fulfill model assumptions (see residual plots and data
transformation in Figure S5 and S6, Table S5). We set the seven
hydrographic basins as a random effect on the intercept of the linear models to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table S6). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). The paths whose
slope had a $P$ value inferior to $0.05$ were reported in the main text, all the
path values are presented in Table S7 and S8).

@@

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 =
.11$, Figure \@ref(fig:sem-plot) B and D). We found that species richness had a
significant negative effect on mean trophic level ($r_\delta = 
`r rich_on_tlvl`$) and connectance ($r_\delta = `r rich_on_ct`$, Figure
\@ref(fig:sem-plot) B and D), meaning that a part of the effects of species richness went 
through network structure.

@@

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Figure \@ref(fig:sem-plot) B, Table
\@ref(tab:indir-sem)), while connectance had no direct effect on total biomass
(Figure \@ref(fig:sem-plot) B).

@@

By its effects on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Figure
\@ref(fig:sem-plot) D).

The total effect of mean trophic level on temporal stability of biomass was of
the same magnitude than the one of species richness (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Figure \@ref(fig:sem-plot) D) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Figure \@ref(fig:sem-plot) D, Table \@ref(tab:indir-sem)).

@@

This study is unique by its level of integration, from environmental
effects to total biomass and its stability, going through species richness and
foodweb structure. It is the first direct assessment of a possible joint effect
of species richness and network structure on both total biomass and its temporal
stability. We found that species richness and mean trophic level had a positive
effect on total biomass. In contrast, species richness had a negative effect on
temporal stability whereas mean trophic level covaried positively. We also found
that the effect of species richness and mean trophic level on total biomass and
its temporal stability were not independent as species richness had a negative
effect on mean trophic level. Furthermore, we found that connectance had no
significant effect on both total biomass and its temporal stability. The results
shows that temporal stability was mediated by changes in species variability 
rather than synchrony. Finally, we found that the magnitude of the effects of
average temperature and altitude was equal or superior to the effects of species
richness and mean trophic level.

@@

Contrasting with previous studies on experimental plant communities
[@grace_integrative_2016], aquatic mesocosms [@yvondurocher_warming_2011;
@dossena_warming_2012] and marine fish communities
[@maureaud_biodiversityecosystem_2019], we found that environmental drivers
overall affect total biomass indirectly rather than directly
[@woods_testing_2020]. First, consistent with research showing that foodwebs in
wider streams support more species and biomass [@allan_stream_2007], we found
that the PCA axis related to stream size had a positive effect on total biomass
through a positive effect on mean trophic level and species richness. Greater
ecosystem size might provide a greater diversity of resources through habitat
heterogeneity, and so higher species richness and total biomass by surface unit,
through resource complementarity. Greater diversity of resources can also
maintain more biomass at high trophic level [@post_ecosystem_2000;
@doi_resource_2009]. Second, the positive effect of the PCA axis related to
temperature on total biomass was of the magnitude of the effects of network
structure and species richness, in line with the mounting body of work reporting
important effects of temperature on the structure and biomass of aquatic systems
[e.g.  @dossena_warming_2012; @edeline_ecological_2013;
@emmrich_geographical_2014; @hattab_forecasting_2016; @woods_testing_2020]. We
found that streams with higher average temperature supported higher diversity as
well as higher average trophic level and total biomass, which is in agreement
with recent findings on streams where the positive effects of temperature were
related to larger nutrient supply [@ogorman_unexpected_2017].

While the PCA axis related with stream size had only a very weak effect on the
temporal stability of total biomass, the PCA axis related to temperature had a
clear negative effect on the temporal stability through a direct destabilizing
effect on populations and through indirect effects. Temperature can have both
stabilizing and destabilizing effects in resource-consumer models depending on
the thermal niche of the species and on how nutrient supply varies with
temperature [@uszko_effects_2017]. In our case, the observed increased
variability of populations might also relate with ecosystem enrichment
[@ogorman_unexpected_2017] because average Biological Oxygen Demand covaries
with temperature in our study. It has been suggested that enrichment can increase CV by
increasing species biomass fluctuations while higher temperature might decrease
decrease stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

Interestingly, the two PCA axes related to environmental conditions had indirect
effects on stability through two conflicting pathways, either stabilising, via
network structure, or destabilising, via species richness. The first pathway
consists in a positive effect of temperature and stream size on mean
trophic level, in line with the idea that wider streams and warmer conditions
are favourable to mesotrophic predators [@allan_stream_2007;
@emmrich_geographical_2014], which in turn stabilises biomass fluctuations,
possibly due to increased top-down regulation. The second pathway consists in a
positive effect of temperature and stream size on species richness, which itself
has a destabilising effect. This result is interesting in that both
pathways had not been evidenced to jointly play, to the best of our knowledge.
Furthermore, it is unclear which pathway is dominating, and under which circumstances,
opening an interesting avenue for further investigations.

@@

# References

>>>>

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in foodwebs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, the presence of species at higher trophic
levels increases total primary production [@wang_biodiversity_2018]. The
addition of predators has also cascading effects on both species variability and
synchrony at lower trophic levels [@teng_dynamics_2004;
@shanafelt_stability_2018]. Overall, these theoretical studies show that the
effects of species richness and network structure, such as connectance and
average trophic level, are potentially conflicting or synergistic in such a way
that they need to be addressed jointly.

@@

Such approach based on data from natural communities can, in addition, give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@oliver_heterogeneous_2010; @franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2017;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017;
@tabi_warming_2019]. Environment also directly affect species richness and
foodweb structure. For instance, larger ecosystems host foodwebs with higher
diversity and trophic levels [@mchugh_dual_2010] while temperature can affect
foodweb diversity and structure in contrasted ways depending on the ecosystem
studied [@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing
data from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electric fishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). The stations where the stream
was of intermediate size could be sampled alternatively with one of the two
sampling protocols. In this case, we kept only the most frequently used sampling
protocol. The sampling season spanned from late spring to autumn. There was low
heterogeneity of the sampling month within stations, with the median sampled
month being mid-August while the median of the standard deviation of the sampled
month within station was $0.7$.

We selected the stations that had been sampled ten years or more. As the concept
of community variability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only
stations with no temporal trends in total biomass were selected (detailed in
appendix B, part 2). However, the results were robust to the relaxation of this
hypothesis (Fig. S7). These selection steps resulted in 99 stations distributed
over seven hydrographic basins (Fig. \@ref(fig:matmet) A).

@@

The body size of all fish individuals was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Species biomass and total biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, species biomass and
total biomass were corrected by the surface sampled (range `r range_surface[1]`-
`r range_surface[2]` square metres), and expressed in grams by square meter
($gm^{-2}$). In the analysis, the total biomass of a site has been defined as
the median of the total biomass across sampling years.

@@

Species richness in a given station has been defined as the total number of species present across
all the sampling events. We controlled species richness for sampling effort by
dividing the total number of species by the total surface sampled in the given
station. Species richness was then expressed in number of species per square
meter sampled.

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Fig.
\@ref(fig:matmet) B), then (2) determine network of a given community by
extracting the metaweb according to the trophic species present at a given
site and sampling event (Fig. \@ref(fig:matmet) C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, alvins feed on plankton while adults
may feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to
a trophic species (Fig. S2). Additionally to the trophic species, we added
seven resource nodes present in the ontogenic food diet database: detritus,
biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos
(grey nodes in Fig. \@ref(fig:matmet) B, C & D). The trophic species and the
resources constituted the 412 nodes of the metaweb, i.e. 45 fish species divided
into nine body size classes plus seven resource nodes. The inference of trophic
interactions was shown to be more accurate when considering body size classes,
i.e. trophic species, than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. Species identity allowed to
differentiate between piscivorous and non piscivorous species, to know which
resources the trophic species are feeding on and to know the minimum size at
which a piscivorous species begin to feed on fishes. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by a body size
range (Fig. S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic specie was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint was included in the
predation window (Fig. S2). The fish-resource trophic links were set between
the trophic species and the resources nodes according the food diet of their
size class, for both the piscivorous and non piscivorous trophic species.
Lastly, the trophic links among resource nodes were set according to the
literature [@allan_stream_2007; @hart_handbook_2008].

@@

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of water temperature, flow, Biological Oxygen Demand (BOD), river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. S4). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, DBO and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

Structural Equation Modeling enables to assess multivariate relationships
between variables. It is a promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and indirect, between variables
[@grace_structural_2008; @grace_causal_2014; @lefcheck_piecewisesem_2016;
@grace_integrative_2016]. We defined a structural equation model explaining the
average annual total biomass as a function of the two synthetic environmental
variables, species richness and foodweb structure as measured by connectance and
average trophic level [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and network
structure directly affect total biomass (Fig.
\@ref(fig:sem-plot) A)
and that species richness affected foodweb structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Fig.
\@ref(fig:sem-plot) C). Instead of the annual total biomass, we modelled the
synchrony and population variability which in turn determined the temporal
stability of biomass.

The structural equation models were based on Gaussian linear models. All variables
except the connectance and average trophic level
were logged to fulfill model assumptions (see residual plots and data
transformation in Fig. S5 and S6, Table S5). We set the seven
hydrographic basins as a random effect on the intercept of the linear models to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table S6). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). The paths whose
slope had a $P$ value inferior to $0.05$ were reported in the main text, all the
path values are presented in Table S7 and S8).

@@

The structural equation models explained a large part of the variation
of species richness, connectance, synchrony, $\overline{CV_{sp}}$ and total
biomass (resp. $R^2 = .43$, $.38$, $.46$, $.47$ and $.32$) but less so for mean
trophic level ($R^2 =
.11$, Fig. \@ref(fig:sem-plot) B and D). We found that species richness had a
significant negative effect on mean trophic level ($r_\delta = 
`r rich_on_tlvl`$) and connectance ($r_\delta = `r rich_on_ct`$, Fig.
\@ref(fig:sem-plot) B and D), meaning that a part of the effects of species richness went 
through network structure.

@@

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and 
$r_\delta = `r tlvl_on_bm`$ resp., Fig. \@ref(fig:sem-plot) B, Table
\@ref(tab:indir-sem)), while connectance had no direct effect on total biomass
(Fig. \@ref(fig:sem-plot) B).

@@

By its effects on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). This effect resulted from two opposite effects. On
the one hand, species richness had a direct negative effect on synchrony
($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Fig.
\@ref(fig:sem-plot) D).

The total effect of mean trophic level on temporal stability of biomass was of
the same magnitude than the one of species richness (resp.
$`r tlvl_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Fig. \@ref(fig:sem-plot) D) and thereby had a total positive
effect on temporal stability ($r_\delta = `r tlvl_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct effect on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Fig. \@ref(fig:sem-plot) D, Table \@ref(tab:indir-sem)).

@@

---
#This study is unique by its level of integration, from environmental
#effects to total biomass and its stability, going through species richness and
#foodweb structure. It is the first direct assessment of a possible joint effect
#of species richness and network structure on both total biomass and its temporal
#stability. We found that species richness and mean trophic level had a positive
#effect on total biomass. In contrast, species richness had a negative effect on
#temporal stability whereas mean trophic level covaried positively. We also found
#that the effect of species richness and mean trophic level on total biomass and
#its temporal stability were not independent as species richness had a negative
#effect on mean trophic level. Furthermore, we found that connectance had no
#significant effect on both total biomass and its temporal stability. The results
#shows that temporal stability was mediated by changes in species variability 
#rather than synchrony. Finally, we found that the magnitude of the effects of
#average temperature and altitude was equal or superior to the effects of species
#richness and mean trophic level. 
---

@@

We found that environmental drivers overall affect total biomass indirectly
[@woods_testing_2020] rather than directly , contrasting with previous studies
[@yvondurocher_warming_2011; @dossena_warming_2012; @grace_integrative_2016;
@maureaud_biodiversityecosystem_2019]. First, consistent with research showing
that foodwebs in wider streams support more species and biomass
[@allan_stream_2007], we found that the PCA axis related to stream size had a
positive effect on total biomass through a positive effect on mean trophic level
and species richness. Greater ecosystem size might provide a greater diversity
of resources through habitat heterogeneity, and so higher species richness and
total biomass by surface unit, through resource complementarity. Greater
diversity of resources can also maintain more biomass at high trophic level
[@post_ecosystem_2000; @doi_resource_2009]. Second, the positive effect of the
PCA axis related to temperature on total biomass was of the magnitude of the
effects of network structure and species richness, in line with the mounting
body of work reporting important effects of temperature on the structure and
biomass of aquatic systems [e.g.  @dossena_warming_2012;
@edeline_ecological_2013; @emmrich_geographical_2014; @hattab_forecasting_2016;
@woods_testing_2020]. We found that warmer streams
supported higher diversity as well as higher average trophic level and total
biomass, which is in agreement with recent findings on streams where the
positive effects of temperature were related to larger nutrient supply
[@ogorman_unexpected_2017].

While the PCA axis related with stream size had only a very weak effect on the
temporal stability of total biomass, the PCA axis related to temperature had a
negative effect on the temporal stability through a direct destabilizing
effect on populations and through indirect effects. Temperature can have both
stabilizing and destabilizing effects in resource-consumer models depending on
the thermal niche of the species and on how nutrient supply varies with
temperature [@uszko_effects_2017]. In our case, the observed increased
variability of populations might also relate with ecosystem enrichment
[@ogorman_unexpected_2017] because average Biological Oxygen Demand covaries
with temperature in our study. Enrichment can increase CV by
increasing species biomass fluctuations while higher temperature might decrease
 stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
#Interestingly, the two PCA axes related to environmental conditions had indirect
#effects on stability through two conflicting pathways, either stabilising, via
#network structure, or destabilising, via species richness. The first pathway
#consists in a positive effect of temperature and stream size on mean
#trophic level, in line with the idea that wider streams and warmer conditions
#are favourable to mesotrophic predators [@allan_stream_2007;
#@emmrich_geographical_2014], which in turn stabilises biomass fluctuations,
#possibly due to increased top-down regulation. The second pathway consists in a
#positive effect of temperature and stream size on species richness, which itself
#has a destabilising effect. This result is interesting in that both
#pathways had not been evidenced to jointly play, to the best of our knowledge.
#Furthermore, it is unclear which pathway is dominating, and under which circumstances,
#opening an interesting avenue for further investigations.
---

@@

# References {-}

<div id="refs"></div>

\clearpage

# Figures and Table {-}
